General purpose gene annotation using deep learning

### 2020-04-07 paper/main.tex
minor
paper

\begin{methods}
\section{Methods}
We worked with 192 animal and 60 plant genomes. The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format. We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation). Both data groups were used seperately during training as well as during evaluation. At the start of our work we split both genome groups into training and test sets. The test sets contained 19 or 6 genomes, respectively. The test genomes were chosen to be of the best possible quality (?) based on collected metadata while also representing a broad phylogenetic spread. Some of the remaining genomes were used for training purposes and all were used for evaluating our models.

@@

Once a set of training genomes was selected, we generated a validation set from it by splitting of 20\% of the FASTA sequences above and below the N90 of each genome, ensuring a proper distribution of large and short sequences in both the training and validation set and a split on the chromosome level. The validation set was only used to roughly judge the performance of our models during training. Evaluation of the annotations on all training and validation genomes was done regularly, as we ultimately only cared about the performance on for the model fully unknown genomes.

@@

\begin{table}[!t]
\processtable{Experimental Results and Comparison\label{Tab:03}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Augustus & 0.632 (+-0.089) & 0.749 (+-0.161)\\
Dilated CNN Baseline & - (+-) & - (+-)\\
Best Helixer model & - (+-) & - (+-) \\
Best model + varied input length & - (+-) & - \\
Best model (+ varied input length) + overlapping &  0.85 (+-0.068) & 0.843 (+-0.151) \\
(Varied input length +) overlapping + ensembling & 0.868 (+-0.064) & 0.863 (+-0.148) \\
\botrule
\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group. Varied input length was only used in the animal case.}
\end{table}

>>>>

\begin{methods}
\section{Methods}
	We worked with 192 animal and 60 plant genomes. The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format. We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation). Both data groups were used seperately during training as well as during evaluation. At the start of our work we split both genome groups into training and test sets. The test sets contained 19 or 6 genomes, respectively. The test genomes were chosen to be of the best possible quality (?) based on collected metadata while also representing a broad phylogenetic spread. Some of the remaining genomes were used for training purposes and all were used for evaluating our models. We call the genomes that were ultimatily used during training the {\it training genomes} and the remaining non-test genomes the {\it validation genomes}.

@@

Once a set of training genomes was selected, we split it up again into a {\it training and validation set}. This split of the training genomes was done to get a quicker sense of the generalization capabilities and was used primarily during training. We split of the validation set by selecting 20\% of the FASTA sequences above and below the N90 of each training genome, ensuring a proper distribution of large and short sequences in both sets and a split on the chromosome level. Evaluation of the annotations on all training and validation genomes was done regularly after a promising model candidate was found based on its validation set performance. Ultimately we only cared about the cross-species performance of a model.

@@

\begin{table}[!t]
\processtable{Experimental Results and Comparison\label{Tab:03}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Augustus & 0.632 (+-0.089) & 0.749 (+-0.161)\\
Dilated CNN Baseline & - (+-) & - (+-)\\
Best Helixer model & - (+-) & - (+-) \\
Best model + varied input length & - (+-) & - \\
Best model (+ varied input length) + overlapping &  0.85 (+-0.068) & 0.843 (+-0.151) \\
(Varied input length +) overlapping + ensembling & 0.868 (+-0.064) & 0.863 (+-0.148) \\
\botrule
\end{tabular}}{Values are the median in Subgenic F1 across all validation and test genomes of the respective group. Varied input length was only used in the animal case. The best model was chosen out of the eight models of the ensemble for having the best performance in the reported setup.}
\end{table}
